Seasonality in governing dynamics at submesoscales in the Kuroshio Extension

### 2016-06-27 writeup/rocha_etal.tex
Minor corrections.

\begin{abstract}
    Two new high-resolution numerical simulations with embedded tides suggest a
    strong modulation in the governing near-surface dynamics at submesoscales
    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late Winter mixed layers
    are prone to baroclinic instabilities, and submesoscale turbulence
    prevails in late Winter/early Spring. Inertia-gravity waves are enhanced near
    the surface when the upper ocean is strongly stratified. Thus the summertime
    re-stratification weakens submesoscale turbulence and enhances intertia-gravity
     waves near the surface. In the Kuroshio Extension region,
    inertia-gravity waves strongly dominate the surface kinetic energy and
    sea-surface height variance in late Summer/early Fall at scales as large as 80 km.
\end{abstract}

@@

\section{Statistics of surface lateral velocity tensor}
To study the seasonality in the surface velocity, we calculate the lateral velocity tensor
\begin{equation}
\nabla_h \bf{u}_h = \left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right] \,.
\end{equation}
The components of the lateral velocity tensor are calculated using a centered
second-order finite differences scheme. We then calculate
the vertical vorticity $\zeta \equiv v_x - u_y$, lateral rate of strain
$\alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}$, and horizontal divergence $\delta \equiv u_x + v_y$.
These diagnostics highlight the submesoscale structures in the flow
\citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.

@@

\begin{acknowledgments}
We thank the MITgcm community and our colleagues at the NASA Advanced
Supercomputing (NAS) Division for their awesome support.
This research was funded by NASA and NSF.
The altimeter products were produced by Ssalto/Duacs
and distributed by Aviso, with support from Cnes (\texttt{http://www.aviso.altimetry.fr/duacs/})
\end{acknowledgments}

@@

\subsection*{How well does the LLC simulations captures high-frequency modes?}
To assess how well the LLC simulations represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

>>>>

\begin{abstract}
    Two new high-resolution numerical simulations with embedded tides suggest a
    strong modulation in the governing near-surface dynamics at submesoscales
    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late Winter mixed layers
    are prone to baroclinic instabilities, and submesoscale turbulence
    prevails in late Winter/early Spring. Inertia-gravity waves are enhanced near
    the surface when the upper ocean is strongly stratified. Thus the summertime
    re-stratification weakens submesoscale turbulence and enhances intertia-gravity
     waves near the surface. In the Kuroshio Extension,
    inertia-gravity waves strongly dominate the surface kinetic energy and
    sea-surface height variance in late Summer/early Fall at submesoscales.
\end{abstract}

@@

\section{Statistics of the surface lateral velocity tensor}
To study the seasonality in the surface velocity, we calculate the lateral velocity tensor
\begin{equation}
\nabla_h \bf{u}_h = \left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right] \,.
\end{equation}
The components of the lateral velocity tensor are calculated using a centered
second-order finite differences scheme. We then calculate
the vertical vorticity $\zeta \equiv v_x - u_y$, lateral rate of strain
$\alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}$, and horizontal divergence $\delta \equiv u_x + v_y$.
These diagnostics highlight the submesoscale structures in the flow
\citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.

@@

\begin{acknowledgments}
We thank the MITgcm community and our colleagues at the NASA Advanced
Supercomputing (NAS) Division for their awesome support.
This research was funded by NASA and NSF.
The altimeter products were produced by Ssalto/Duacs
and distributed by Aviso, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
\end{acknowledgments}

@@

\subsection*{How well do the LLC simulations captures high-frequency modes?}
To assess how well the LLC simulations represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.
